---
title: "tenta_svar"
author: "Isak Berntsson"
date: "8/15/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(geoR)
#'install.packages("extraDistr")'
library(extraDistr)


source("ExamData.R")
set.seed(1337)
```

# Tenta Juni 2022

## Question 1, Demand of a product

### 1.b
```{r}

alpha_n = 2326
beta_n = 7

ndraws = 1E4
theta_draws = rgamma(ndraws,shape = alpha_n, rate=beta_n)
q_draws = rpois(ndraws,theta_draws)


prob = mean(q_draws>350)
hist(q_draws, main="histogram of draws", sub= paste("P(Q_6 > 350) = ",prob))
abline(v=350, col="red")



```

### 1.c

```{r}

grid = seq(0,1000,by=1)
utils = matrix(nrow=length(grid),ncol=1)

for (i in 1:length(grid)){
  utils[i] = mean(utility_func(grid[i],q_draws))
}

best_a = grid[which.max(utils)]
plot(x=grid, y = utils, sub=paste("best a = ",best_a))
abline(v=best_a)
```

## Question 2, Regression

### 2.a

 

```{r}
num_covs = dim(X)[2]
num_draws = 10000

#priors

mu_0 = rep(0,num_covs)
omega_0 = 1/(100) * diag(6)
v_0 = 1
sigma2_0 = 100^2

res = BayesLinReg(y,X, mu_0, omega_0, v_0, sigma2_0, num_draws)

betas = res$betaSample

for (i in 1:num_covs){
  sample = betas[,i]
  sample_interval = quantile(sample, probs=c(0.005,0.995))
  sample_mean = mean(sample)
  
  hist(sample, main = colnames(X)[i],
       sub = paste(" \n low: ",sample_interval[1],"\n high:",sample_interval[2], "\n mean = ", sample_mean, sep=""), xlab="'")
  
  abline(v=sample_mean, col="blue")
  abline(v=sample_interval, col="red")
}


```
B_1 has a significant positive value according to our analysis. The exact value likely(99%) between [5.9, 15.8]


### 2.b

```{r}

sigma_sample = res$sigma2Sample

posterior_mean = mean(sigma_sample**.5)
posterior_median = median(sigma_sample**.5)

list(posterior_mean = mean(sigma_sample**.5),
posterior_median = median(sigma_sample**.5))
```

### 2.c


```{r}

x1_lo = min(X[,2])
x1_hi = max(X[,2])
x1_grid = seq(x1_lo,x1_hi,by=.1)

X_test = matrix(nrow=length(x1_grid),ncol = num_covs)

colnames(X_test) = colnames(X)
X_test[,1] = 1
X_test[,2]=x1_grid
X_test[,3] = x1_grid ** 2
X_test[,4] = 27
X_test[,5] = 27** 2
X_test[,6] = 27 * x1_grid
head(X_test)


for (i in 1:length(x1_grid)){
  x_i = X_test[i,]
  
  mu_i = betas %*% (x_i)
  
  interval = quantile(mu_i, probs=c(0.025,0.975))
  if (i ==1){
    plot(x=x_i[2], y=interval[1],xlim=c(x1_lo-1,x1_hi+1), ylim=c(0,500), xlab="x1", ylab = "mu", main = "95% ETCI of mu, depending on x1")
    points(x=x_i[2], y=interval[2])
  }else{  
    points(x=x_i[2],y=interval[1])
    points(x=x_i[2], y=interval[2])
  }
}


points(X[,2], y, col="red")
```

### 2.d

The effect from x1 depends on x2 only in the sixth Feature. x1*X2.

Consider the two extreme cases: 
  1. x2 = n, n -> 0
  2. x2 = M, M -> inf

in case one: the dependence is incredibly small and the sixth feature is inconsequential.

in case two: the dependence is very large and the total effect from x1 depends almost totally on x2. 

In general, there will be some dependence and our effect from x1 is dependent on x2.
Additionally, we see that the 95% interval for the effect is negative.


```{r}
plot(density(betas[,6]))

quantile(betas[,6],probs=c(0.025,0.975))
```

## 2.e



```{r}
x1 = 50
x2 = 25

x_test = c(1, x1, x1^2, x2, x2^2, x1*x2)
```


```{r}
preds = betas %*% x_test + rnorm(length(sigma_sample),0,sd=sqrt(sigma_sample))

hist(preds)





```
## 2.f

```{r}
leny = length(y)
maxy = max(y)


res = betas %*% t(X)
head(res)


#for every draw of betas and sigma
#perform prediction for all data points
#select highest

#compare this set of highest predictions against highest measured.
# pval = P(max(predicted) > max(measured)), should be around 50%

pred_max = matrix(nrow= ndraws,ncol=1)
for (i in 1:ndraws){
  
  preds = rnorm(leny,res[i,],sqrt(sigma_sample[i]))
  pred_max[i] = max(preds)
}
head(pred)
prob = mean(pred_max > maxy)

hist(pred_max, sub=paste("pval",prob))
abline(v=max(y), col="red")
abline(v=median(pred_max),col="red")


```
```{r}
y=cars$dist

#print(var(y))
sample = rinvchisq(1E5,length(y)-1,var(y))

plot(density(sample))
abline(v=var(y))
```


